---
title: "Homework 3"
author: "Calvin Wong"
date: "10/1/2018"
output: html_document
---

```{r}
library(utils)
library(dplyr)
library(ggplot2)
library(DATA606)
library(psych)

```


3.2 Area under the curve, Part II. What percent of a standard normal distribution N(μ =
0, ! = 1) is found in each region? Be sure to draw a graph.

(a) Z > −1.13 
87.1%
```{r}
normalPlot(mean = 0, sd = 1, bounds = c(-1.13, 4))
```

(b) Z < 0.18 
57.1%
```{r}
normalPlot(mean = 0, sd = 1, bounds = c(-4, 0.18))
```

(c) Z > 8 
-> 0%
```{r}
normalPlot(mean = 0, sd = 1, bounds = c(8, Inf))
```

(d) |Z| < 0.5
13.4%
```{r}
normalPlot(mean = 0, sd = 1, bounds = c(1 - round(pnorm(.5, mean = 0, sd = 1),digits=3), round(pnorm(.5, mean = 0, sd = 1),digits=3)))
```


3.4 Triathlon times, Part I. In triathlons, it is common for racers to be placed into age and gender groups. Friends Leo and Mary both completed the Hermosa Beach Triathlon, where Leo competed in the Men, Ages 30 - 34 group while Mary competed in the Women, Ages 25 - 29 group. Leo completed the race in 1:22:28 (4948 seconds), while Mary completed the race in 1:31:53 (5513 seconds). Obviously Leo finished faster, but they are curious about how they did within their respective groups. Can you help them? Here is some information on the performance of their groups:
• The finishing times of the Men, Ages 30 - 34 group has a mean of 4313 seconds with a standard deviation of 583 seconds.
• The finishing times of the Women, Ages 25 - 29 group has a mean of 5261 seconds with a standard deviation of 807 seconds.
• The distributions of finishing times for both groups are approximately Normal.
Remember: a better performance corresponds to a faster finish.
(a) Write down the short-hand for these two normal distributions.
Men 30 - 34, N(mean = 4313, sd = 583)
Women 25 - 29, N(mean = 5261, sd = 807)

(b) What are the Z-scores for Leo’s and Mary’s finishing times? What do these Z-scores tell you?
Leo Z = (4948 - 4313) / 583 = 1.089
Mary Z = (5513 - 5261) / 807 = 0.312

(c) Did Leo or Mary rank better in their respective groups? Explain your reasoning.
Leo ranked better in his group because his time is associated with a higher percentile within his group.

(d) What percent of the triathletes did Leo finish faster than in his group?
86.2%
```{r}
normalPlot(mean = 0, sd = 1, bounds = c(-4, 1.089))
```

(e) What percent of the triathletes did Mary finish faster than in her group?
62.2%
```{r}
normalPlot(mean = 0, sd = 1, bounds = c(-4, 0.312))
```

(f) If the distributions of finishing times are not nearly normal, would your answers to parts (b) - (e) change? Explain your reasoning.
b) would not change. However for parts (c) to (e), we cannot use the normal probability table to calculate probabilities and percentile without a normal model.


3.18 Heights of female college students. Below are heights of 25 female college students.
(a) The mean height is 61.52 inches with a standard deviation of 4.58 inches. Use this information to determine if the heights approximately follow the 68-95-99.7% Rule.

Height follow the 68-95-99.7% rule very closely at 68-96-100%.
```{r}
height <- c(54,55,56,56,57,58,58,59,60,60,60,61,61,62,62,63,63,63,64,65,65,67,67,69,73)
describe(height)
length(height[abs(height - mean(height)) < sd(height)])/length(height) ## at 1 sd
length(height[abs(height - mean(height)) < 2*sd(height)])/length(height)
length(height[abs(height - mean(height)) < 3*sd(height)])/length(height)
```

(b) Do these data appear to follow a normal distribution? Explain your reasoning using the graphs provided below.

The distribution is unimodal and symmetric.The superimposed normal curve seems to approximate the distribution pretty well. The points on the normal probability plot also seem to follow a straight line. There is a possible outlier on the higher end that is apparent in both graphs, but it is not too extreme. We can say that the distribution is nearly normal.

3.22 Defective rate. A machine that produces a special type of transistor (a component of computers) has a 2% defective rate. The production is considered a random process where each transistor is independent of the others.
(a) What is the probability that the 10th transistor produced is the first with a defect?

The probability that the 10th transistor produced is the first with a defect is 1.67%
```{r}
(1-.02)^9 * .02
```

(b) What is the probability that the machine produces no defective transistors in a batch of 100?

The probability that the machine produces no defective transistors in a batch of 100 is 13.26%

```{r}
(1-.02)^100
```

(c) On average, how many transistors would you expect to be produced before the first with a defect? What is the standard deviation?

50 transistors produced before the first with a defect. The sd is 49.50.

```{r}
1/.02
sqrt((1-.02)/(.02^2))
```


(d) Another machine that also produces transistors has a 5% defective rate where each transistor is produced independent of the others. On average how many transistors would you expect to be produced with this machine before the first with a defect? What is the standard deviation?

20 transistors produced before the first with a defect. The sd is 19.49.

```{r}
1/.05
sqrt((1-.05)/(.05^2))
```

(e) Based on your answers to parts (c) and (d), how does increasing the probability of an event a↵ect the mean and standard deviation of the wait time until success?

Yes it does. When the probability is lower there is a higher chance of success.

3.38 Male children. While it is often assumed that the probabilities of having a boy or a girl are the same, the actual probability of having a boy is slightly higher at 0.51. Suppose a couple plans to have 3 kids.
(a) Use the binomial model to calculate the probability that two of them will be boys.

```{r}
dbinom(2, 3, 0.51)
```

(b) Write out all possible orderings of 3 children, 2 of whom are boys. Use these scenarios to calculate the same probability from part (a) but using the addition rule for disjoint outcomes.
Confirm that your answers from parts (a) and (b) match.

BBG, BGB, GBB

The answers match.

```{r}
(0.51^2) * 0.49 * 3
```

(c) If we wanted to calculate the probability that a couple who plans to have 8 kids will have 3 boys, briefly describe why the approach from part (b) would be more tedious than the approach from part (a).

The approach from part (b) will be more tedious because it requires drawing out all the scenarios for 8 combinations.

3.42 Serving in volleyball. A not-so-skilled volleyball player has a 15% chance of making the serve, which involves hitting the ball so it passes over the net on a trajectory such that it will land in the opposing team’s court. Suppose that her serves are independent of each other.
(a) What is the probability that on the 10th try she will make her 3rd successful serve?

3.9%

```{r}
n <- 10    
k <- 3     
p <- 0.15  
factorial(n-1) / (factorial(k-1) * (factorial(n - k))) * p^k * (1-p)^(n-k)
```

(b) Suppose she has made two successful serves in nine attempts. What is the probability that her 10th serve will be successful?

The probability that her 10th serve will be successful is 0.15 since each serve is independant.

(c) Even though parts (a) and (b) discuss the same scenario, the probabilities you calculated should be different. Can you explain the reason for this discrepancy?

The probabilities calculated will be different because part (a) refers to the probability that the 10th try will yield the 3rd successful serve, wherelse, (b) refers to independent probability outcomes.
